Estimating Fixed-frame Galaxy Magnitudes in the SDSS Spectroscopic Survey

### 2002-02-15 docs/paper/paper.tex
plot tweak
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@44 75dff457-8751-0410-ab1a-cd701654ce6b

% Authorship determined by those I was directly involved with
% in performing this work as well as those responsible for the photo-z
% plates and those who have characterized the filter curves (since
% these might be publicly distributed). 
\author{
Michael R. Blanton\altaffilmark{\ref{NYU}},
%Tamas Budavari\altaffilmark{\ref{JHU}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
Istv\'an Csabai\altaffilmark{\ref{JHU}},
Mamoru Doi\altaffilmark{\ref{Tokyo}},
Daniel Eisenstein\altaffilmark{\ref{Arizona}},
James E. Gunn\altaffilmark{\ref{Princeton}}, and
David W. Hogg\altaffilmark{\ref{NYU}}
%David J. Schlegel\altaffilmark{\ref{Princeton}}
%Julianne Dalcanton\altaffilmark{\ref{UW}},
%Jon Loveday\altaffilmark{\ref{Sussex}},
%Michael A. Strauss\altaffilmark{\ref{Princeton}},
%Mark SubbaRao\altaffilmark{\ref{Chicago}},
%David H. Weinberg\altaffilmark{\ref{Ohio}},
%John E. Anderson, Jr.\altaffilmark{\ref{Fermilab}},
%James Annis\altaffilmark{\ref{Fermilab}},
%Neta A. Bahcall\altaffilmark{\ref{Princeton}},
%Mariangela Bernardi\altaffilmark{\ref{Chicago}},
%J. Brinkmann\altaffilmark{\ref{APO}},
%Robert J. Brunner\altaffilmark{\ref{Caltech}},
%Scott Burles\altaffilmark{\ref{Fermilab}},
%Larry Carey\altaffilmark{\ref{UW}},
%Francisco J. Castander\altaffilmark{\ref{Chicago}, \ref{Pyrenees}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Douglas Finkbeiner\altaffilmark{\ref{Berkeley}},
%Scott Friedman\altaffilmark{\ref{JHU}},
%Joshua A. Frieman\altaffilmark{\ref{Fermilab}},
%Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
%G. S. Hennessy\altaffilmark{\ref{USNO}},
%Robert B. Hindsley\altaffilmark{\ref{USNO}},
%Takashi Ichikawa\altaffilmark{\ref{Tokyo}},
%\v{Z}eljko Ivezi\'{c}\altaffilmark{\ref{Princeton}},
%Stephen Kent\altaffilmark{\ref{Fermilab}},
%G. R.~Knapp\altaffilmark{\ref{Princeton}},
%D. Q.~Lamb\altaffilmark{\ref{Chicago}},
%R. French Leger\altaffilmark{\ref{UW}},
%Daniel C. Long\altaffilmark{\ref{APO}},
%Robert H. Lupton\altaffilmark{\ref{Princeton}},
%Timothy A.~McKay\altaffilmark{\ref{Michigan}},
%Avery Meiksin\altaffilmark{\ref{Edinburgh}},
%Aronne Merelli\altaffilmark{\ref{Caltech}},
%Jeffrey A. Munn\altaffilmark{\ref{USNO}},
%Vijay Narayanan\altaffilmark{\ref{Princeton}},
%Matt Newcomb\altaffilmark{\ref{CarnegieMellon}},
%R. C. Nichol\altaffilmark{\ref{CarnegieMellon}},
%Sadanori Okamura\altaffilmark{\ref{Tokyo}},
%Russell Owen\altaffilmark{\ref{UW}},
%Jeffrey R.~Pier\altaffilmark{\ref{USNO}},
%Adrian Pope\altaffilmark{\ref{JHU}},
%Marc Postman\altaffilmark{\ref{STScI}},
%Thomas Quinn\altaffilmark{\ref{UW}},
%Constance M. Rockosi\altaffilmark{\ref{Chicago}},
%Donald P. Schneider\altaffilmark{\ref{PennState}}, 
%Kazuhiro Shimasaku\altaffilmark{\ref{Tokyo}},
%Walter A. Siegmund\altaffilmark{\ref{UW}},
%Stephen Smee\altaffilmark{\ref{Maryland}},
%Yehuda Snir\altaffilmark{\ref{CarnegieMellon}},
%Chris Stoughton\altaffilmark{\ref{Fermilab}},
%Christopher Stubbs\altaffilmark{\ref{UW}},
%Alexander S.~Szalay\altaffilmark{\ref{JHU}},
%Gyula P.~Szokoly\altaffilmark{\ref{Potsdam}},
%Aniruddha R.~Thakar\altaffilmark{\ref{JHU}},
%Christy Tremonti\altaffilmark{\ref{JHU}},
%Douglas L. Tucker\altaffilmark{\ref{Fermilab}},
%Alan Uomoto\altaffilmark{\ref{JHU}},
%Dan vanden Berk\altaffilmark{\ref{Fermilab}},
%Michael S. Vogeley\altaffilmark{\ref{Drexel}},
%Patrick Waddell\altaffilmark{\ref{UW}},
%Brian Yanny\altaffilmark{\ref{Fermilab}},
%Naoki Yasuda\altaffilmark{\ref{NAOJ}},
%and Donald G.~York\altaffilmark{\ref{Chicago}}
}

@@

The reader may ask why, if we are basing our $K$-corrections on a full
model of the galaxy SED, we do not correct to bandpasses with simpler
shapes (say, Gaussians). The reason is that we want the effective
bandpasses to actually have been observed for some galaxies in the
sample, so that the $K$-correction for those galaxies is formally and
actually zero. This property is desirable because, as noted above,
there are uncertainties in the $K$-corrections. Nevertheless, we note
that synthesizing such magnitudes may be appropriate in certain
situations, having the virtue (similarly to the $AB$ magnitude system)
that such magnitudes are very easy to comprehend and synthesize from
theory. For example, if we wanted to convert our results into a form
easily comprehended by the future astronomers discussed in the first
paragraph of Section \ref{motivation}, it would be best to calculate
fluxes within wide top-hat filters so that the difficulties of
understanding the magnitude system were minimized.

@@

The SDSS (\citealt{york00a}) is producing imaging and spectroscopic
surveys over $\pi$ steradians in the Northern Galactic Cap. A
dedicated 2.5m telescope (Siegmund {\it et al.}, in preparation) at
Apache Point Observatory, Sunspot, New Mexico, images the sky in five
bands between 3000 and 10000 \AA\ ($u$, $g$, $r$, $i$, $z$;
\citealt{fukugita96a}) using a drift-scanning, mosaic CCD camera
(\citealt{gunn98a}), detecting objects to a flux limit of $r'\sim
22.5$. The ultimate goal is to spectroscopically observe 900,000
galaxies, (down to $r_{\mathrm{lim}}'\approx 17.65$), 100,000 Luminous
Red Galaxies (LRGs; Eisenstein {\it et al.}, in preparation), and
100,000 QSOs (\citealt{fan99a}; Newberg {\it et al.}, in preparation).
This spectroscopic follow up uses two digital spectrographs on the
same telescope as the imaging camera. Many of the details of the
galaxy survey are described in a description of the galaxy target
selection paper (\citealt{strauss02a}). Other aspects of the survey
are described in the Early Data Release (EDR;
\citealt{stoughton01a}). The survey has begun in earnest, and has
obtained about 20\% of its intended data.

As of February 2002, the SDSS had targeted and taken spectra of
approximately XXX galaxies over XXX square degrees. As described
below, only a subset of these galaxies are used here to develop the
templates, though we calculate the SED fits for all of them. The
results of the photometric pipeline for all of these objects were
extracted from the SDSS Operational Database {\bf ref munn?}. The
photometry used here for the bulk of these objects was the same as
that used when the objects were targeted. However, for those objects
which were in the EDR photometric catalog, we used the better
calibrations and photometry from the EDR. The versions of the SDSS
pipeline PHOTO used for the reductions of these data ranged from {\tt
v5\_0} to {\tt v5\_2}. The changes in the treatment of relatively
small galaxies, which accounts for most of our sample, did not change
substantially throughout these versions.

@@

The response functions $R(\lambda)$ have been measured using a
monochrometer by {\bf how ref Doi}.  Using a model for the atmospheric
transmission and the reflectivity of the primary and secondary
mirrors, one can then model the response of the entire system. For
each bandpass in the SDSS, there are six different CCDs; it has been
shown that the differences between these six camera columns are
small. The resulting set of filter curves is shown in Figure
\ref{response_sdss}, in comparison to a model of a galaxy spectral
energy distribution observed at $z=0$ (a 4 Gyr old instantaneous burst
from the models of \citealt{bruzual93a}).

We extract one-dimensional spectra from the two-dimension images using
a pipeline created specifically for the SDSS instrumentation
(\citealt{schlegel02a}), which also fits for the redshift of each
spectrum. The official SDSS redshifts are obtained from a different
pipeline (\citealt{subbarao02a}). The two independent versions provide
a consistency check on the redshift determination. They are consistent
(for galaxies) at over the 99\% level.\footnote{The official pipeline
tends to work better for objects with unusual spectra, such as certain
types of stars and QSOs.}.

We use two types of magnitudes determined by the SDSS. First, the
modified form of the magnitude described by \citet{petrosian76a}, as
described in \citet{strauss02a}.  These magnitudes have a nearly
constant metric aperture as a function redshift; in addition, all the
bands use the same aperture, so the measured SED corresponds (to
within the effects of seeing) to the SED of an identifiable region of
the galaxy. However, for faint objects, the Petrosian magnitudes tend
to become noisy. Thus, for galaxies with $z>0.25$ we instead use the
higher signal-to-noise ``model magnitudes.'' Model magnitudes are
calculated in all bands using a single weighted aperture. The weighted
aperture is the better fitting model (pure exponential or pure de
Vaucouleurs) to the galaxy image in the $r$-band. In
\citet{stoughton01a} we are explicitly warned (in paragraphs written
by the lead author of this paper!)  not to mix Petrosian and model
magnitudes in a single analysis, since they measure galaxies in very
different ways. Nevertheless, for the purposes of contraining galaxy
template SEDs, it is perfectly acceptable to use {\it any}
well-defined part of any galaxy.  For our purposes, it is more
important to have high signal-to-noise measurement for any
well-defined part of our galaxies, even if the region of the galaxies
probed and their total estimated fluxes are inconsistent with
redshift. We extinction-correct both types of magnitudes using the
dust maps of \citet{schlegel98a}.

@@

For the SED space (our $\vv{v}_i(\lambda)$), we use the subspace
defined by linear combinations of ten Bruzual-Charlot instantaneous
burst models with ages ranging from $3 \times 10^7$ to $2\times
10^{10}$ years, five with metallicity $Z=0.02$ and five with
$Z=0.004$, all assuming a Salpeter Initial Mass Function. We find that
we cannot recover information about the SED below a certain scale,
except perhaps around the 4000\AA\ break. Thus, we smooth most of the
wavelength regime of the templates using a Gaussian with a standard
deviation of $\sigma = 300$\AA; around 4000\AA, we smooth only with
$\sigma = 150$\AA.

@@

%The resulting four templates are shown in Figure \ref{k_espec_plot} in
%the range $2000<\lambda<12000$\AA, the region which is constrained by
%our observations. 
The top two panels of Figure \ref{k_coeffdist_plot} show the pairwise
joint distributions of $a_1/a_0$, $a_2/a_0$, and $a_3/a_0$. In the
bottom panel, three spectra taken from a one-dimensional sequence
along the galay locus are shown, showing that the spectra become
progressively bluer along that sequence. Note that constraints on the
SED become poor at the bluest and reddest edges of this diagram (the
peak of the $z$-band response is at 8700\AA). Therefore, the odd
behavior near the edges should not be taken too seriously.

@@

An important test of the reasonableness of the fits is to check that
for a fixed type of galaxy, the distribution of the fixed frame colors
depends little on redshift. Because the galaxies shown in Figures
\ref{k_coeffdist_plot} and \ref{k_model_plot} are inhomogeneously
selected --- some are Main Sample galaxies, some are Luminous Red
Galaxies, and some are selected from the photometric redshift plates
--- we will split our sample into two well-defined sets of
objects. First, we choose a set of Main Sample galaxies in the
luminosity range $-21.5 < M_{\band{0.1}{r}} < -21$. We reconstruct the
four colors $^{0.1}(u-g)$, $^{0.1}(g-r)$, $^{0.1}(r-i)$, and
$^{0.1}(i-z)$ for these galaxies and plot the reconstructions as a
function of redshift in left-hand panels of Figure
\ref{main_colors_plot}. The plots show a general consistency in the
fixed-frame color distribution of these objects with redshift. The
right-hand panels show the distribution for redshifts $z<0.13$ (solid
histogram) and $z>0.13$ (dotted histogram) . Small changes are
discernible in the distributions, mostly attributable to the increased
errors at higher redshift. Note that for $z<0.1$ the $^{0.1}(u-g)$
color depends on an extrapolation of the SED in the blue, while for
$z>0.1$ the $^{0.1}(i-z)$ color depends on an extrapolation of the SED
in the red.

Second, we choose a set of Cut I LRGs with $-22.8 < M_{\band{0.1}{r}}
< -22.5$. Now we reconstruct the colors to a fixed-frame corresponding
to the rest-frame bands used to observe a galaxy at $z=0.3$.  Near
$z=0.3$ there is some singular behavior of the colors, which we will
discuss in more detail later. Again, for $z<0.3$ the $^{0.3}(u-g)$
color relies on an extrapolation, which is probably responsible for
the scatter in that color at low redshifts. Otherwise, the
distribution of colors is remarkably constant. $^{0.3}(g-r)$ (which is
very similar to $^{0.0}(u-g)$) experiences a blueward shift of about
0.1 magnitudes between $z=0.1$ and $z=0.45$, which is attributable to
passive evolution.

In short, these fits to galaxy SEDs provide estimates which reproduce
the galaxy photometry nearly to the level of the errors in the
photometry itself, seem physically reasonable, and are consistent over
the range of redshifts we consider ($0.0 < z < 0.5$). The only
remaining problem is the aforementioned difficulties at $z=0.3$,
discussed in more detail in the next section.

@@

It is clear from Figure \ref{k_kcorrect_plot} that $z=0.3$ is a
special redshift regime where our fits become unphysical. In fact, it
is easy to demonstrate that the diagonal spur (going from the upper
left to the lower right) around the red dot in the upper right panel
of Figure \ref{k_coeffdist_plot} is due to galaxies at $z\sim
0.3$. The difficulty arises because the 4000 \AA\ break lies in the
gap between $g$ and $r$ at $z=0.3$. The algorithm requires strong
gradients in the spectrum in this area to explain the 4000\AA break
--- however, the gap in the bands causes a degeneracy arise.

To demonstrate this degeneracy, Figure \ref{spur} shows the spectrum
corresponding to the direction of the spur in the $a_1/a_0$-$a_3/a_0$
plane of Figure \ref{k_coeffdist_plot}. We overplot the \band{0.3}{g}
and \band{0.3}{r} bands. The spur has a strong peak in the gap between
the two bands at $z=0.3$. Because of this gap, at $z\sim 0.3$ the
algorithm tends to use this direction to help fit the $u$, $i$, and
$z$ fluxes, with relatively little harm done to the $g$ and $r$
fits. This explains why the reconstructed fluxes in Figure
\ref{k_model_plot} fit well even at $z=0.3$, even though the spur at
$z=0.3$ in the distribution of coefficients is so evident.  This
difficulty is why, in Section \ref{data}, we exclude the regime around
$z\sim 0.3$; otherwise the templates became distracted by the
degeneracies in this regime.

The question arises of what to do about this problem. One solution is
to simply work around it. It is clear we cannot use the fits presented
here to $K$-correct galaxies which are at $z=0.3$ to other redshifts
reliably. However, nothing prevents one from $K$-correcting galaxies
at other redshifts {\it to} $z=0.3$. If you are considering the LRG
sample, the median redshift is close to $z=0.3$ anyway, so this is a
viable solution (and the one adopted in Figure \ref{lrg_colors_plot}).

A second solution, not adopted here, is to place stronger constraints
in the SED fit on the form of the 4000 \AA\ break. One obvious way of
doing this is to use the spectra to constrain the size of the break
(making it a weak constraint to prevent biases). This constraint could
be applied in a linear fashion to retain the speed of our method ---
except for the fact that you would have to read in the spectra to
apply the $K$-correction. For simplicity, and because the aforemention
work-around is so easy to obey, we have not implemented this solution
here.

\section{$K$-corrections}
\label{kcorrection}

\subsection{Testing the $K$-corrections}

We show, in Figure \ref{kcorrect}, the $K$-corrections to $z=0.1$
inferred from this method, for all five SDSS bands. Note that the
$K$-corrections are largest (and thus most uncertain) in
$\band{0.1}{u}$ and $\band{0.1}{g}$.

>>>>

% Authorship determined by those I was directly involved with
% in performing this work as well as those responsible for the photo-z
% plates and those who have characterized the filter curves (since
% these might be publicly distributed). 
\author{
Michael R. Blanton\altaffilmark{\ref{NYU}},
%Tamas Budavari\altaffilmark{\ref{JHU}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
Istv\'an Csabai\altaffilmark{\ref{JHU}},
Mamoru Doi\altaffilmark{\ref{Tokyo}},
Daniel Eisenstein\altaffilmark{\ref{Arizona}},
James E. Gunn\altaffilmark{\ref{Princeton}}, and
David W. Hogg\altaffilmark{\ref{NYU}}
David J. Schlegel\altaffilmark{\ref{Princeton}}
%Julianne Dalcanton\altaffilmark{\ref{UW}},
%Jon Loveday\altaffilmark{\ref{Sussex}},
%Michael A. Strauss\altaffilmark{\ref{Princeton}},
%Mark SubbaRao\altaffilmark{\ref{Chicago}},
%David H. Weinberg\altaffilmark{\ref{Ohio}},
%John E. Anderson, Jr.\altaffilmark{\ref{Fermilab}},
%James Annis\altaffilmark{\ref{Fermilab}},
%Neta A. Bahcall\altaffilmark{\ref{Princeton}},
%Mariangela Bernardi\altaffilmark{\ref{Chicago}},
%J. Brinkmann\altaffilmark{\ref{APO}},
%Robert J. Brunner\altaffilmark{\ref{Caltech}},
%Scott Burles\altaffilmark{\ref{Fermilab}},
%Larry Carey\altaffilmark{\ref{UW}},
%Francisco J. Castander\altaffilmark{\ref{Chicago}, \ref{Pyrenees}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Douglas Finkbeiner\altaffilmark{\ref{Berkeley}},
%Scott Friedman\altaffilmark{\ref{JHU}},
%Joshua A. Frieman\altaffilmark{\ref{Fermilab}},
%Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
%G. S. Hennessy\altaffilmark{\ref{USNO}},
%Robert B. Hindsley\altaffilmark{\ref{USNO}},
%Takashi Ichikawa\altaffilmark{\ref{Tokyo}},
%\v{Z}eljko Ivezi\'{c}\altaffilmark{\ref{Princeton}},
%Stephen Kent\altaffilmark{\ref{Fermilab}},
%G. R.~Knapp\altaffilmark{\ref{Princeton}},
%D. Q.~Lamb\altaffilmark{\ref{Chicago}},
%R. French Leger\altaffilmark{\ref{UW}},
%Daniel C. Long\altaffilmark{\ref{APO}},
%Robert H. Lupton\altaffilmark{\ref{Princeton}},
%Timothy A.~McKay\altaffilmark{\ref{Michigan}},
%Avery Meiksin\altaffilmark{\ref{Edinburgh}},
%Aronne Merelli\altaffilmark{\ref{Caltech}},
%Jeffrey A. Munn\altaffilmark{\ref{USNO}},
%Vijay Narayanan\altaffilmark{\ref{Princeton}},
%Matt Newcomb\altaffilmark{\ref{CarnegieMellon}},
%R. C. Nichol\altaffilmark{\ref{CarnegieMellon}},
%Sadanori Okamura\altaffilmark{\ref{Tokyo}},
%Russell Owen\altaffilmark{\ref{UW}},
%Jeffrey R.~Pier\altaffilmark{\ref{USNO}},
%Adrian Pope\altaffilmark{\ref{JHU}},
%Marc Postman\altaffilmark{\ref{STScI}},
%Thomas Quinn\altaffilmark{\ref{UW}},
%Constance M. Rockosi\altaffilmark{\ref{Chicago}},
%Donald P. Schneider\altaffilmark{\ref{PennState}}, 
%Kazuhiro Shimasaku\altaffilmark{\ref{Tokyo}},
%Walter A. Siegmund\altaffilmark{\ref{UW}},
%Stephen Smee\altaffilmark{\ref{Maryland}},
%Yehuda Snir\altaffilmark{\ref{CarnegieMellon}},
%Chris Stoughton\altaffilmark{\ref{Fermilab}},
%Christopher Stubbs\altaffilmark{\ref{UW}},
%Alexander S.~Szalay\altaffilmark{\ref{JHU}},
%Gyula P.~Szokoly\altaffilmark{\ref{Potsdam}},
%Aniruddha R.~Thakar\altaffilmark{\ref{JHU}},
%Christy Tremonti\altaffilmark{\ref{JHU}},
%Douglas L. Tucker\altaffilmark{\ref{Fermilab}},
%Alan Uomoto\altaffilmark{\ref{JHU}},
%Dan vanden Berk\altaffilmark{\ref{Fermilab}},
%Michael S. Vogeley\altaffilmark{\ref{Drexel}},
%Patrick Waddell\altaffilmark{\ref{UW}},
%Brian Yanny\altaffilmark{\ref{Fermilab}},
%Naoki Yasuda\altaffilmark{\ref{NAOJ}},
%and Donald G.~York\altaffilmark{\ref{Chicago}}
}

@@

The reader may ask why, if we are basing our $K$-corrections on a full
model of the galaxy SED, we do not correct to bandpasses with simpler
shapes (say, Gaussians). The reason is that we want the effective
bandpasses to actually have been observed for some galaxies in the
sample, so that the $K$-correction for those galaxies is formally and
actually zero. This property is desirable because, as noted above,
there are uncertainties in the $K$-corrections. Nevertheless, we note
that synthesizing such magnitudes may be appropriate in certain
situations, as they have the virtue (similarly to the $AB$ magnitude
system) that such magnitudes are very easy to comprehend and
synthesize from theory. For example, if we wanted to convert our
results into a form easily comprehended by the hypothetical future
astronomers discussed in the first paragraph of Section
\ref{motivation}, it would be best to calculate fluxes within wide
top-hat filters so that the difficulties of understanding the
magnitude system were minimized.

@@

The SDSS (\citealt{york00a}) is producing imaging and spectroscopic
surveys over $\pi$ steradians in the Northern Galactic Cap. A
dedicated 2.5m telescope (Siegmund {\it et al.}, in preparation) at
Apache Point Observatory, Sunspot, New Mexico, images the sky in five
bands between 3000 and 10000 \AA\ ($u$, $g$, $r$, $i$, $z$;
\citealt{fukugita96a}) using a drift-scanning, mosaic CCD camera
(\citealt{gunn98a}), detecting objects to a flux limit of $r'\sim
22.5$. The ultimate goal is to spectroscopically observe 900,000
galaxies, (down to $r_{\mathrm{lim}}'\approx 17.65$), 100,000 Luminous
Red Galaxies (LRGs; Eisenstein {\it et al.}, in preparation), and
100,000 QSOs (\citealt{fan99a}; Newberg {\it et al.}, in preparation).
This spectroscopic follow up uses two digital spectrographs on the
same telescope as the imaging camera. Many of the details of the
galaxy survey are described in a description of the galaxy target
selection paper (\citealt{strauss02a}). Other aspects of the survey
are described in the Early Data Release (EDR;
\citealt{stoughton01a}). The survey has begun in earnest, and has
obtained about 20\% of its intended data.  As of February 2002, the
SDSS had targeted and taken spectra of approximately XXX galaxies over
XXX square degrees.

As described below, only a subset of these galaxies are used here to
develop the templates, though we calculate the SED fits for all of
them. The results of the photometric pipeline for all of these objects
were extracted from the SDSS Operational Database {\bf ref munn?}. The
photometry used here for the bulk of these objects was the same as
that used when the objects were targeted. However, for those objects
which were in the EDR photometric catalog, we used the better
calibrations and photometry from the EDR. The versions of the SDSS
pipeline PHOTO used for the reductions of these data ranged from {\tt
v5\_0} to {\tt v5\_2}. The changes in the treatment of relatively
small galaxies, which accounts for most of our sample, did not change
substantially throughout these versions.

@@

The response functions of the filter-CCD combinations have been
measured using a monochrometer by {\bf how ref Doi}.  Using a model
for the atmospheric transmission and the reflectivity of the primary
and secondary mirrors, one can then model the response of the entire
system. For each bandpass in the SDSS, there are six different CCDs;
it has been shown that the differences between these six camera
columns are small; what differences exist are largely fit out during
the survey calibration. The resulting set of filter curves is shown in
Figure \ref{response_sdss}, in comparison to a model of a galaxy
spectral energy distribution observed at $z=0$ (a 4 Gyr old
instantaneous burst from the models of \citealt{bruzual93a}).

The results of the spectroscopic observations are treated as follows.
We extract one-dimensional spectra from the two-dimension images using
a pipeline ({\tt specBS v4\_8}) created specifically for the SDSS
instrumentation (\citealt{schlegel02a}), which also fits for the
redshift of each spectrum. The official SDSS redshifts are obtained
from a different pipeline (\citealt{subbarao02a}). The two independent
versions provide a consistency check on the redshift
determination. They are consistent (for galaxies) at over the 99\%
level.\footnote{The official pipeline tends to work better for objects
with unusual spectra, such as certain types of stars and QSOs.}.

We use two types of magnitudes determined by the SDSS. First, the
modified form of the magnitude described by \citet{petrosian76a}, as
described in \citet{strauss02a}.  These magnitudes have a nearly
constant metric aperture as a function redshift; in addition, all the
bands use the same aperture, so the measured SED corresponds (to
within the effects of seeing) to the SED of an identifiable region of
the galaxy. However, for faint objects, the Petrosian magnitudes tend
to become noisy. Thus, for galaxies with $z>0.25$ we instead use the
higher signal-to-noise ``model magnitudes.'' Model magnitudes are
calculated in all bands using a single weighted aperture; thus, the
measured colors again correspond to an identifiable region of the
galaxy, though a different one than would be measured by the Petrosian
magnitude. The weighted aperture is the better fitting model (pure
exponential or pure de Vaucouleurs) to the galaxy image in the
$r$-band. In \citet{stoughton01a} we are explicitly warned (in a
paragraph written by the lead author of this paper!)  not to mix
Petrosian and model magnitudes in a single analysis, since they
measure galaxies in very different ways. Nevertheless, for the
purposes of contraining galaxy template SEDs, it is perfectly
acceptable to use {\it any} well-defined part of any galaxy.  For our
purposes, it is more important to have high signal-to-noise
measurements of galaxy colors than to probe exactly the same regions
of galaxies at low and high redshift.  We extinction-correct both
types of magnitudes using the dust maps of \citet{schlegel98a}. {\tt
describe fiber mags}

@@

For the SED space (our $\vv{v}_i(\lambda)$), we use the subspace
defined by linear combinations of ten Bruzual-Charlot instantaneous
burst models with ages ranging from $3 \times 10^7$ to $2\times
10^{10}$ years, five with metallicity $Z=0.02$ and five with
$Z=0.004$, all assuming a Salpeter Initial Mass Function. Since we
cannot recover information about the SED below a certain scale, we
smooth most of the wavelength regime of the templates using a Gaussian
with a standard deviation of $\sigma = 300$\AA; for the region of the
spectrum which contains the sharpest gradients, around 4000\AA, we
smooth only with $\sigma = 150$\AA. {\bf not clear anymore that this
is the best thing to do; see the g/r gap discusssion}

@@

%The resulting four templates are shown in Figure \ref{k_espec_plot} in
%the range $2000<\lambda<12000$\AA, the region which is constrained by
%our observations. 
The top two panels of Figure \ref{k_coeffdist_plot} show the pairwise
joint distributions of $a_1/a_0$, $a_2/a_0$, and $a_3/a_0$ for a
random subset of about 10,000 of the galaxies in the SDSS sample (not
only the ones which we used to fit the templates). In the bottom
panel, three spectra taken from a one-dimensional sequence along the
galay locus are shown, showing that the spectra become progressively
bluer along that sequence. Note that constraints on the SED become
poor at the bluest and reddest edges of this diagram (the peak of the
$z$-band response is at 8700\AA). Therefore, the odd behavior near the
edges should not be taken too seriously. In addition, note the spur
extending from upper left to lower right through the left edge of the
($a3/a0$)-($a1/a0$) plane. Most of the objects in this extendend spur
are at around $z=0.3$. We discuss in Section \ref{grgap} why this spur
arises and what to do about it.

@@

An important test of the reasonableness of the fits is to check that
for a fixed type of galaxy, the distribution of the fixed frame colors
depends only weakly on redshift. Because the galaxies shown in Figures
\ref{k_coeffdist_plot} and \ref{k_model_plot} are inhomogeneously
selected --- some are Main Sample galaxies, some are Luminous Red
Galaxies, and some are selected from the photometric redshift plates
--- we will split our sample into two well-defined sets of objects to
perform this test. First, we choose a set of Main Sample galaxies in
the luminosity range $-21.5 < M_{\band{0.1}{r}} < -21$. We reconstruct
the four colors $^{0.1}(u-g)$, $^{0.1}(g-r)$, $^{0.1}(r-i)$, and
$^{0.1}(i-z)$ for these galaxies and plot the reconstructions as a
function of redshift in left-hand panels of Figure
\ref{main_colors_plot}. The plots show a general consistency in the
fixed-frame color distribution of these objects with redshift. The
right-hand panels show the distribution for redshifts $z<0.13$ (solid
histogram) and $z>0.13$ (dotted histogram) . Small changes are
discernible in the distributions, mostly attributable to the increased
errors at higher redshift. Note that for $z<0.1$ the $^{0.1}(u-g)$
color depends on an extrapolation of the SED in the blue, while for
$z>0.1$ the $^{0.1}(i-z)$ color depends on an extrapolation of the SED
in the red.

Second, we choose a set of Cut I LRGs with $-22.8 < M_{\band{0.3}{r}}
< -22.5$. For this sample we reconstruct the colors at $z=0.3$.  Near
$z=0.3$ there is some singular behavior of the colors, which we will
discuss in more detail in Section \ref{grgap}. Again, for $z<0.3$ the
$^{0.3}(u-g)$ color relies on an extrapolation, which is probably
responsible for the scatter in that color at low redshifts. Otherwise,
the distribution of colors is remarkably constant. $^{0.3}(g-r)$
(which is very similar to $^{0.0}(u-g)$) experiences a blueward shift
of about 0.1 magnitudes between $z=0.1$ and $z=0.45$, which is
attributable to passive evolution.

In short, these fits to galaxy SEDs provide estimates which reproduce
the galaxy photometry nearly to the level of the errors in the
photometry itself, seem physically reasonable, and are consistent over
the range of redshifts we consider ($0.0 < z < 0.5$). The only
remaining problem is the aforementioned difficulties at $z=0.3$,
discussed in more detail in the next subsection.

@@

It is clear from Figure \ref{k_kcorrect_plot} that $z=0.3$ is a
special redshift regime where our fits become unphysical. In fact, it
is easy to demonstrate that the diagonal spur (going from the upper
left to the lower right) around the red dot in the upper right panel
of Figure \ref{k_coeffdist_plot} is due almost completely to galaxies
at $z\sim 0.3$. The spur arises because the 4000 \AA\ break lies in
the gap between $g$ and $r$ at $z=0.3$, visible in Figure
\ref{response_sdss}. Our algorithm has strong gradients in the
spectrum around the 4000\AA\ break, which is necessary to properly
model galaxy spectra. In the presence of these sharp gradients, the
gap in the bands causes certain combinations of the template spectra
to be degenerate at around $z=0.3$.

To demonstrate this degeneracy, Figure \ref{spur} shows the spectrum
corresponding to the direction of the spur in the
$(a_1/a_0)$-$(a_3/a_0)$ plane of Figure \ref{k_coeffdist_plot}; this
direction is represented by the vector $-0.6 \vv{e}_1(\lambda) +
\vv{e}_3$. We overplot the \band{0.3}{g} and \band{0.3}{r} bands. The
spur has a strong peak in the gap between the two bands at
$z=0.3$. Because of this gap, at $z\sim 0.3$ the algorithm tends to a
component in this direction to help fit the $u$, $i$, and $z$ fluxes,
with relatively little extra $\chi^2$ contribution from the $g$ and
$r$ observations. Thus the reconstructed fluxes in Figure
\ref{k_model_plot} fit well even at $z=0.3$, even though the spur at
$z=0.3$ in the distribution of coefficients is so evident.  This
difficulty is why, in Section \ref{data}, we exclude the regime around
$z\sim 0.3$; otherwise the templates became distracted by the
degeneracies in this regime.

The question arises of what to do about this problem. One solution is
to simply work around it. It is clear we cannot use the fits presented
here to $K$-correct galaxies which are at $z=0.3$ to other redshifts
reliably --- their coefficients $a_k$ are unusual and probably
unphysical. However, nothing prevents one from $K$-correcting galaxies
at other redshifts {\it to} $z=0.3$. If you are considering the LRG
sample, the median redshift is close to $z=0.3$ anyway, so this is a
viable solution (and the one adopted in Figure \ref{lrg_colors_plot}).

A second solution is to place extra constraints on the fit which
prevent the spur in the coefficients. One can imagine fitting the
distribution in the $(a_1/a_0)$-$(a_3/a_0)$ plane with a simple model,
such as a multigaussian fit, and using the resulting fit as a prior in
one's fit (with gaussians, the extra constraints would be linear and
thus easily accounted for). One could design these constraints to be
strong enough (at least near the redshifts in question) to prevent the
spur. However, the disadvantage of this solution is that for some
scientific projects the change in the galaxy SED is exactly what one
is trying to measure. In these cases, it may be inappropriate to be
putting explicit {\it a priori} constraints on exactly what one is
trying to measure.

A third solution, not adopted here, is to place stronger constraints
in the SED fit on the form of the 4000 \AA\ break. One obvious way of
doing this is to use the spectra to constrain the size of the break
(making it a weak constraint to prevent biases). This constraint could
be applied in a linear fashion to retain the speed of our method ---
except for the fact that you would have to read in the spectra to
apply the $K$-correction. For simplicity, and because the aforemention
work-around is so easy to obey, we have not implemented this solution
here.

\subsection{$K$-corrections in the SDSS}

We show, in Figure \ref{k_kcorrect_plot}, the resulting $K$-corrections to
$z=0.3$ inferred from this method, for all five SDSS bands. Note that
the $K$-corrections are largest (and thus most uncertain) in
$\band{0.3}{u}$ and $\band{0.3}{g}$. Remember that the $K$-corrections
to $\band{0.3}{u}$ are extrapolations for $z<0.3$ and that the
$K$-corrections to $\band{0.3}{z}$ are extrapolations for $z>0.3$, so
those results should not be taken too seriously. (Although the
$K$-corrections given are not {\it too} unreasonable).
